---
title: "Project 3"
author: "Pruthvi Ranjan Reddy Pati"
date: "3/3/2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(nnet)
library(tidyr)
library(ggplot2)
```

## Data Preparation

```{r c1,warning=FALSE,error=FALSE}

FAA_1 <- read_excel("FAA1(1).xls")
FAA_2 <- read_excel("FAA2(1).xls")
FAA_2 <- mutate(FAA_2, duration = NA)
FAA <- rbind(FAA_1, FAA_2)
FAA <- FAA[,c(1,3,4,6,7,8,2,5)]
FAA <- FAA %>% 
  filter(duration > 40 | is.na(duration)) %>% 
  filter(between(speed_ground, 30, 140))  %>%  
  filter(between(speed_air, 30, 140) | is.na(speed_air))  %>% 
  filter(height > 6)
FAA <- arrange_(FAA, .dots = colnames(FAA)[1:6])
FAA <- FAA[!duplicated(FAA[,1:6]),]
FAA <- FAA %>% 
  mutate(aircraft_01 = case_when(
    FAA$aircraft == "airbus" ~ 0,
    TRUE ~ 1))
FAA$aircraft_01 <- as.factor(FAA$aircraft_01)
faa_multi<-FAA
faa_multi$y<-ifelse(faa_multi$distance<1000,1,
              ifelse(faa_multi$distance>=2500,3,2))
faa_multi<-faa_multi[,-c(1,6)]
faa_multi$y<-as.factor(faa_multi$y)
summary(faa_multi)
```
##Multinomial Regression
```{r c2,warning=FALSE,error=FALSE}
multi_reg<-multinom(formula= y ~ duration+aircraft_01+no_pasg+ speed_ground+height+pitch, data=faa_multi)


summary(multi_reg)

```

* Performing step regression after mean imputing `duration`
```{r c3,warning=FALSE,error=FALSE}

faa_multi$duration<-ifelse(is.na(faa_multi$duration),mean(faa_multi$duration,na.rm=TRUE),faa_multi$duration)

multi_reg1<-multinom(formula= y ~ duration+aircraft_01+no_pasg+ speed_ground+height+pitch, data=faa_multi)

step_multi<-step(multi_reg1)
```

##Final Model
```{r final,warning=FALSE,error=FALSE}
summary(step_multi)

```

##Plots
```{r c4,warning=FALSE,error=FALSE}
sg<-mutate(faa_multi, speed=cut_number(speed_ground,5)) %>%
group_by(speed, y) %>% summarise(count=n()) %>%
group_by(speed) %>% mutate(etotal=sum(count),
proportion=count/etotal)
ggplot(sg, aes(x=speed, y=proportion, group=y,linetype=y)) + geom_line()

h<-mutate(faa_multi, height=cut_number(height,5)) %>%
group_by(height, y) %>% summarise(count=n()) %>%
group_by(height) %>% mutate(etotal=sum(count),
proportion=count/etotal)
ggplot(h, aes(x=height, y=proportion, group=y,linetype=y)) + geom_line()

p<-mutate(faa_multi, pitch=cut_number(pitch,5)) %>%
group_by(pitch, y) %>% summarise(count=n()) %>%
group_by(pitch) %>% mutate(etotal=sum(count),
proportion=count/etotal)
ggplot(p, aes(x=pitch, y=proportion, group=y,linetype=y)) + geom_line()

air<-group_by(faa_multi,aircraft_01, y) %>%
summarise(count=n()) %>% group_by(aircraft_01) %>%
mutate(etotal=sum(count), proportion=count/etotal)
ggplot(air, aes(x=aircraft_01, y=proportion,
group=y, linetype=y)) + geom_line()

```

##Observation for the FAA agent:
* `Aircraft type`,`speed ground`,`height`,`pitch` are the most influential on level of the landing distance
* From the above plots we can observe,the aircrafty type`Boeing` has higher chances of either having very high landing distances or very low compared to `Airbus`.Also, the probabilty ratio of landing distance of level 3 and 1 will multiply by e^9 factor
* With `pitch` increasing the chances of risky landing increases and also obviously with `speed ground`
 




##Prediction

```{r c5,warning=FALSE,error=FALSE}
preds<-
data.frame(faa_multi$y,predict(step_multi,data.frame(faa_multi),type="probs"))
xtabs(~predict(step_multi)+faa_multi$y)



```
#2


* Number of Passengers is a non-negative integer and is a count. Poisson Distribution might be good choice to explain Number of Passengers. 

* For Poisson regression , original distance variable is used

```{r c6,warning=FALSE,error=FALSE}
poisson_data<-FAA[,-c(1,6)]
poisson_data$duration<-ifelse(is.na(poisson_data$duration),mean(poisson_data$duration,na.rm=TRUE),poisson_data$duration)

pois_reg<-glm(no_pasg~.,family = "poisson",data=poisson_data[,-6])

step_pois<-step(pois_reg)

summary(step_pois)


```
##Observations:
* Poisson regression is performed.
* From the step wise regression it can be observed that there is a only small variation in the AIC of different model.
* `Height` can be considered the most significant of all the variables to predict `Number of Passengers`



